\LARGE \bf Cascading Situation Assessment for Robots

### 2017-02-28 paper.tex
Update on Overleaf.

    We introduce \uwds, a novel lightweight framework for \emph{cascading
    spatio-temporal situation assessment} in robotics. \uwds allows programmers
    to represent the robot's environment as real-time distributed data
    structures, containing both scene graphs (for representation of 3D
    geometries) and timelines (for representation of temporal events). \uwds
    supports \emph{cascading} representations: the environment is viewed as a
    set of \emph{worlds} that can each have different spatial and temporal
    granularities, and may depend on each others.  \uwds also provides a set of
    high-level client libraries and tools to introspect and manipulate the
    environment models.

@@

\uwds is a distributed and lightweight open-source
framework\footnote{\url{https://github.com/severin-lemaignan/underworlds}} that
enables robot programmers to build and refine in real-time spatial and temporal
models of the environment surrounding a robot. \uwds makes it possible to share
these world models amongst the software components running on the robot.
Besides, \uwds permits to represent and manipulate in a distributed manner
\emph{multiple alternatives} to the current, perceived world model. For
instance, the world with some objects filtered out; the world `viewed' from the
perspective of another agent; an hypothetical world resulting from the simulated
application of a plan, etc.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{overview}
    \caption{Schema of a possible \uwds network: eight (user-written \&
    architecture specific) \emph{clients} (in blue) are sharing environment
    models through four independent \emph{worlds} (made of joint spatial and
    temporal models). This architecture enables successive and modular
    refinement of the models (\emph{cascading} situation assessment),
    effectively adapted to each client's needs. \textit{Dashed yellow nodes represent
    other possible components in the system that do not interact directly with
    the \uwds network}.}

@@

\uwds is specifically inspired by the geometric and temporal reasoner
{\sc Spark} (\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.
{\sc Spark} acts as a situation assessment reasoner that generates symbolic
knowledge from the geometry of the environment with respect to relations between
objects, robots and humans. It also takes into account the different perspective
that each agent has on the environment. {\sc Spark} embeds a
modality-independent geometric model of the environment that serves both as
basis for the fusion of the perception modalities and as bridge with the
symbolic layer~\cite{lemaignan2016artificial}. This geometric model is built
from 3D CAD models of the objects, furnitures and robots, and full body, rigged
models of humans.  It is updated at run-time by the robot's sensors.
Likewise, \uwds embeds a grounded amodal model of the environment, updated
online from the robot's sensors (sensor fusion).

However, {\sc Spark} is a monolithic module, that does not support sharing its
internal 3D model to other external components. In contrast, \uwds focuses on
offering a shared and distributed representation of the environment within the
robot's software architecture.

Work on distributed scene graphs~\cite{naef2003blue} has been previously applied
to robotics to provide a shared 3D representation of the robot's environment
(for instance, the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or the
\emph{Deep State Representation} proposed in~\cite{bustos2016unified}).  \uwds
offers a similar distribution mechanism for 3D scene graphs and extends it to
temporal representations. Besides, \uwds further extends this line of work by
providing the ability to create, manipulate and share \emph{multiple alternative
worlds}, corresponding for instance to filtered or hypothetical \emph{views} on
the initial, perceived model of the environment.

@@

These components exhibit different needs in terms of model,
like different nominal spatial and/or temporal resolutions. For instance, a 3D
motion planner would typically use coarse 3D models of surrounding objects to
lower the computational load while planning, while a module assessing the
visibility of objects might need on the contrary  high-resolution models 
for accurate 3D rendering. This requirement of multiple
task-specific representations has been framed as the need for \emph{deep
representations} by Beetz~\cite{beetz2010towards}.

@@

Traditional robotic middlewares like ROS are not particularly well-suited to
deal with these different needs: full geometric data can be represented, but are
not first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
position are not particularly easy to perform with ROS) and critically,
representing and reasoning on alternative states of the environment is not
directly feasible.

Representing alternative states is however often highly desirable. For instance,
software components manipulating environment models typically perform better if
the models are physically consistent. However, low-level perception inaccuracies
often introduce hard-to-avoid physical inconsistencies (like detected objects
floating in the air, or on the contrary wrongly inset into other objects).
Therefore, a post-process stage (for instance, using a physics simulation
engine) is needed to displace the objects seen by the robot into
physically-correct positions. Implemented with a classical approach (for
instance, using ROS TF frames), we would represent an object {\tt book} with two
frames: the original frame (\eg {\tt book\_frame\_raw}) and a second one
computed by the physics engine (\eg {\tt book\_frame\_corrected}). Such an
approach leads to the robot's 3D model being cluttered with multiple frames and
does not scale well.

Another example pertains to geometric task planning: a geometric task planner
typically needs to reason over hypothetical future states of the environment
(``what happen if I move this glass onto that pile of books?'').  The planner
generates many possible future states, that themselves might require further
processing (for instance, running a physics simulation). Such a tool would
benefit a flexible representation system, where models are derived from each
others, with partial modifications and different timescales.

A third example relates to human-robot interaction scenarios where \emph{perspective
taking} is important (a prototypical example being the game `I spy with my
little eyes', as implemented in~\cite{ros2010which}). Perspective taking is a
cognitive skill that relies on the ability for an agent to project itself in
others' `eyes' to estimate what they see from their perspectives. 
Perspective taking has been previously implemented in robotics by temporary placing
virtual cameras at eyes position for each of the humans tracked by the
robot~\cite{ros2010solving}. While acceptable for simple cases, such an approach
does not permit to maintain truly independent models of the environment for each
agent, and in particular does not permit false-beliefs
modelling~\cite{lemaignan2015mutual} whereas separate, independent world models
would effectively support such a skill.

Lastly, geometric \emph{pre-supposition accommodation} makes another interesting
case for alternative worlds representation. \emph{Pre-supposition accommodation}
comes originally from linguistics, where it describes the mechanism by which
\emph{context is adjusted [...] to accept [...] a sentence that imposes certain
requirements on the context in which it is
processed}~\cite{vonfintel2008presupposition}. In the context of spatio-temporal
representations, we call pre-supposition accommodation the ability for an agent
to adjust its model so that it matches some contextual constraint. For instance,
if A tells B to ``catch the red balloon behind you'', B might create a
representation of an imaginary red balloon, placed behind him, even without
actually observing the balloon: B \emph{accommodates} the \emph{pre-supposition}
of a red balloon being present behind himself. Endowing robots with this
capability has been touched by Mavridis \etal within their multi-modal
\emph{Grounded Situation Model}~\cite{Mavridis2006}. However, to the best of our
knowledge, a general framework which would enable robots to accommodate spatial
and temporal pre-suppositions by deriving imaginary worlds from existing ones
has not been proposed so far.

\uwds addresses this need and the main contribution of this work is a generic
approach to \textbf{represent and share multiple parallel representations of the
world}. \uwds does so by permitting clients to clone existing worlds, modify
them, and re-share them, without the cost of duplicating geometric data (as
explained in section~\ref{design}). By organizing clients in a network
(Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
in a loosely-coupled modular approach to spatio-temporal world representation
that we call \emph{cascading situation assessment}.

@@

Worlds contain both a geometric model and a temporal model. The geometric
model is represented as a scene graph. The scene graph has a unique root node,
to which a tree of other nodes is parented.

@@

Every node has a unique ID, a parent, a 3D transformation relative to the parent
and an optional name. \emph{Object} nodes optionally store as well pointers to
their associated meshes. Importantly, mesh data (or other geometric datasets
like point clouds) are \emph{not} stored within the nodes themselves. \uwds
represents geometric data as immutable data, labelled with their hash
(preventing \textit{de facto} data duplication).  Nodes only store the hash
corresponding to the desired geometric data, and the actual data is pulled from
the server by the clients whenever they actually need it (for rendering for
instance).

@@

Complementing the spatial representation encapsulated in the scene graph, each
world also store the world's \emph{timeline}. This data structure is shared
and synchronized amongst the clients in the same way as the scene graph.
Clients can record in the timeline both \emph{events} (duration-less states) and
\emph{situations}, \ie states with a start time and a (possibly open-ended) end time.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{uwds-screenshot}
    \caption{Screenshot of the {\tt uwds view} 3D visualization and manipulation
    tool. In this particular example, the 3D meshes have been pre-loaded using
    {\tt uwds load}. Their positions are then updated at run time using the
    robot's sensors and proprioception (joint state).}
    \label{fig|uwds-view}
\end{figure}

@@

As previously mentioned, worlds are composite data structures comprising of a
scene graph and a timeline. These data structures are synchronized using
Google's gRPC message passing framework\footnote{\url{http://www.grpc.io/}}, ensuring
high throughput, reliability and cross-platform/cross-language support. The \uwds
API is specifically discussed hereafter, in section~\ref{api}.

\uwds is meant to broadcast complex environment representations (typically
including large geometric datasets, like meshes) in real-time. \uwds itself does
not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor
fusion, physics simulation, etc. -- are performed by the clients themselves) and
as such, the performance bottleneck is essentially the network's data
throughput.  In that regard, one of the simple yet critical optimization
performed by \uwds is automatic caching of mesh data. Mesh data are not
transmitted when nodes are updated; only a hash value of the mesh data. The
client can then request the full data whenever they actually need it.

@@

\subsubsection{3D Visualization and manipulation}

Interestingly, while \uwds deals with 3D geometries and scenes, it does
represent 3D entities purely as data structures: no visual representation is
involved (and as such, the \uwds server and core libraries do not depend on any
graphics library like OpenGL). However, for all practical purposes, the ability
to visualize the content of a scene is desirable. \uwds provides a standard
client, {\tt uwds view}, that performs real-time 3D rendering of worlds,
using OpenGL (Figure~\ref{fig|uwds-view}).

@@

Geometric data like meshes can be loaded into a world (or rather, into the whole
network as geometric data are treated as immutable resources shared amongst all
worlds) by simply sending an array of vertices and faces to the \uwds server.
As such, any client can load 3D data that is automatically made available to the
whole network. For example, an object recognition module recognizes a new
object. It sends the mesh to the \uwds server and creates a new node which
references the mesh. At the next frame update, the 3D viewer sees a new node
referencing a yet unknown mesh. The viewer request the mesh data from the
server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
the node transformation is updated, the 3D viewer can re-render the node without
having to download the mesh from the server again.

@@

Specialized command-line tools are also available to list at run-time the worlds
and their content ({\tt uwds ls}) or to display detailed informations on a
specific node ({\tt uwds show}).

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{l2tor-photo2}
    \caption{A user manipulating objects on top of a RFID mat (providing object
    localization) with a NAO robot, while an operator monitors
    the interaction on a tablet that display an {\tt uwds view} client.}

@@

\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
in order to conceptualize and visualize the spatial relations and visibility of
the physical objects that participants interact with. In one of the scenarios of this
project, participants are instructed by a NAO robot to manipulate a number of
Duplo-like animals on top of a printed background paper which serves as a
landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
pedagogical strategies to promote spatial relation learning by object
manipulation. Such instructions include commands like to place the elephant
\textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside}
the house.  To assess the performance, the robot and the operator need to
compute spatial relation between objects from the visual perspective of the
participant.

@@

The interaction scenario requires a number of \uwds clients. The client are
distributed across two computers (the robot and a Microsoft Windows-operated
tablet) that communicate with each other in order to synchronize the worlds
representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
the \uwds network, that consists in the following clients:

@@

Finally, we are evaluating the use of a hybrid platform that utilizes both virtual and
physical objects at the same time.

>>>>

    We introduce \uwds, a novel lightweight framework for \emph{cascading
    spatio-temporal situation assessment} in robotics. \uwds allows programmers
    to represent the robot's environment as real-time distributed data
    structures, containing both scene graphs (for representation of 3D
    geometries) and timelines (for representation of temporal events). \uwds
    supports \emph{cascading} representations: the environment is viewed as a
    set of \emph{worlds} that can each have different spatial and temporal
    granularities, and may depend on each other.  \uwds also provides a set of
    high-level client libraries and tools to introspect and manipulate the
    environment models.

@@

\uwds is a distributed and lightweight open-source
framework\footnote{\url{https://github.com/severin-lemaignan/underworlds}} that
enables robot programmers to build and refine spatial and temporal
models of the environment surrounding a robot in real-time. \uwds makes it possible to share
these world models amongst the software components running on the robot.
Additionally, \uwds enables users to represent and manipulate
\emph{multiple alternatives} to the current, perceived world model in a distributed manner. For
instance, the world with some objects filtered out; the world `viewed' from the
perspective of another agent; a hypothetical world resulting from the simulated
application of a plan, etc.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{overview}
    \caption{Schema of a possible \uwds network: eight \emph{clients} (user-written \&
    architecture specific; in blue) are sharing environment
    models through four independent \emph{worlds} (made from joint spatial and
    temporal models). This architecture enables successive and modular
    refinement of the models (\emph{cascading} situation assessment),
    effectively adapted to each client's needs. \textit{Dashed yellow nodes represent
    other possible components in the system that do not interact directly with
    the \uwds network}.}

@@

\uwds is specifically inspired by the geometric and temporal reasoner
{\sc Spark} (\emph{SPAtial Reasoning \& Knowledge})~\cite{sisbot2011situation}.
{\sc Spark} acts as a situation assessment reasoner that generates symbolic
knowledge from the geometry of the environment with respect to relations between
objects, robots and humans. It also takes into account the different perspective
that each agent has on the environment. {\sc Spark} embeds a
modality-independent geometric model of the environment that serves both as
basis for the fusion of the perception modalities and as bridge with the
symbolic layer~\cite{lemaignan2016artificial}. This geometric model is built
from 3D CAD models of the objects, furniture and robots, and full body, rigged
models of humans.  It is updated at run-time by the robot's sensors.
Likewise, \uwds embeds a grounded amodal model of the environment, updated
online from the robot's sensors (sensor fusion).

However, {\sc Spark} is a monolithic module, that does not support sharing its
internal 3D model with other external components. In contrast, \uwds focuses on
offering a shared and distributed representation of the environment within the
robot's software architecture.

Work on distributed scene graphs~\cite{naef2003blue} has been previously applied
to robotics to provide a shared 3D representation of the robot's environment
(for instance, the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or the
\emph{Deep State Representation} proposed in~\cite{bustos2016unified}).  \uwds
offers a similar distribution mechanism for 3D scene graphs and extends it to
temporal representations. Besides, \uwds further extends this line of work by
providing the ability to create, manipulate and share \emph{multiple alternative
worlds}. As an example, these could correspond to filtered or hypothetical \emph{views} on
the initial, perceived model of the environment.

@@

These components exhibit different needs in terms of representation,
like different nominal spatial and/or temporal resolutions. For instance, a 3D
motion planner would typically use coarse 3D models of surrounding objects to
lower the computational load while planning, while a module assessing the
visibility of objects might need high-resolution models 
for accurate 3D rendering. This requirement of multiple
task-specific representations has been framed as the need for \emph{deep
representations} by Beetz~\cite{beetz2010towards}.

@@

% JK: like software, the plural of middleware is middleware :)
Traditional robotic middleware, like ROS, is not particularly well-suited to
deal with these different needs: full geometric data can be represented, but are
not first-class citizens (a basic task like displaying a 3D mesh at a dynamic
position is not particularly easy to perform with ROS) and critically,
representing and reasoning on alternative states of the environment is not
directly feasible.

Representing alternative states is however often highly desirable. For instance,
software components manipulating environment models typically perform better if
the models are physically consistent. However, low-level perception inaccuracies
often introduce hard-to-avoid physical inconsistencies (like detected objects
floating in the air, or wrongly inset into other objects).
Therefore, a post-process stage (for instance, using a physics simulation
engine) is needed to displace the objects seen by the robot into
physically-correct positions. Implemented with a classical approach (for
instance, using ROS TF frames), we would represent an object {\tt book} with two
frames: the original frame (\eg {\tt book\_frame\_raw}) and a second one
computed by the physics engine (\eg {\tt book\_frame\_corrected}). Such an
approach leads to the robot's 3D model being cluttered with multiple frames and
does not scale well.

Another example pertains to geometric task planning: a geometric task planner
typically needs to reason over hypothetical future states of the environment
(``what happen if I move this glass onto that pile of books?'').  The planner
generates many possible future states, that in turn might require further
processing (for instance, running a physics simulation). Such a tool would
benefit a flexible representation system, where models are derived from each
other, with partial modifications and different timescales.

A third example relates to human-robot interaction scenarios where \emph{perspective
taking} is important (a prototypical example being the game `I spy with my
little eye', as implemented in~\cite{ros2010which}). Perspective taking is a
cognitive skill that relies on the ability for an agent to project itself in
others' `eyes' to estimate what they see from their perspectives. 
Perspective taking has previously been implemented in robotics by temporarily placing
virtual cameras at eye locations for each of the humans tracked by the
robot~\cite{ros2010solving}. While acceptable for simple cases, such an approach
does not maintain truly independent models of the environment for each
agent, and in particular does not permit false-belief
modelling~\cite{lemaignan2015mutual} whereas separate, independent world models
would effectively support such a skill.

Lastly, geometric \emph{pre-supposition accommodation} makes another interesting
case for alternative worlds representation. \emph{Pre-supposition accommodation}
originally comes from linguistics, where it describes the mechanism by which
\emph{context is adjusted [...] to accept [...] a sentence that imposes certain
requirements on the context in which it is
processed}~\cite{vonfintel2008presupposition}. In the context of spatio-temporal
representations, we call pre-supposition accommodation the ability of an agent
to adjust its model so that it matches some contextual constraint. For instance,
if A tells B to ``catch the red balloon behind you'', B might create a
representation of an imaginary red balloon, placed behind him, even without
actually observing the balloon: B \emph{accommodates} the \emph{pre-supposition}
of a red balloon being present behind himself. Endowing robots with this
capability has been touched upon by Mavridis \etal within their multi-modal
\emph{Grounded Situation Model}~\cite{Mavridis2006}. However, to the best of our
knowledge, a general framework which would enable robots to accommodate spatial
and temporal pre-suppositions by deriving imaginary worlds from existing ones
has not been proposed so far.

\uwds addresses this need and the main contribution of this work is a generic
approach to \textbf{represent and share multiple parallel representations of the
world}. \uwds does so by permitting clients to clone existing worlds, modify
them, and re-share them, without the cost of duplicating geometric data (as
explained in section~\ref{design}). By organising clients in a network
(Figure~\ref{fig|scene}), worlds can be made dependent on each other, resulting
in a loosely-coupled modular approach to spatio-temporal world representation
that we call \emph{cascading situation assessment}.

@@

Worlds contain both a geometric model and a temporal model. The geometric
model is represented as a scene graph. The scene graph has a unique root node,
in which a tree of other nodes is parented.

@@

Every node has a unique ID, a parent, a 3D transformation relative to the parent
and an optional name. \emph{Object} nodes optionally store as well pointers to
their associated meshes. Importantly, mesh data (or other geometric datasets
like point clouds) are \emph{not} stored within the nodes themselves. \uwds
represents geometric data as immutable data, labelled with their hash
(preventing \textit{de facto} data duplication).  Nodes only store the hash
corresponding to the desired geometric data, and the actual data is pulled from
the server by the clients whenever they actually need it (rendering for
instance).

@@

Complementing the spatial representation encapsulated in the scene graph, each
world also store the world's \emph{timeline}. This data structure is shared
and synchronised amongst the clients in the same way as the scene graph.
Clients can record in the timeline both \emph{events} (duration-less states) and
\emph{situations}, \ie states with a start time and a (possibly open-ended) end time.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{uwds-screenshot}
    \caption{Screenshot of the {\tt uwds view} 3D visualisation and manipulation
    tool. In this particular example, the 3D meshes have been pre-loaded using
    {\tt uwds load}. Their positions are then updated at run time using the
    robot's sensors and proprioception (joint state).}
    \label{fig|uwds-view}
\end{figure}

@@

As previously mentioned, worlds are composite data structures comprising of a
scene graph and a timeline. These data structures are synchronised using
Google's gRPC message passing framework\footnote{\url{http://www.grpc.io/}}, ensuring
high throughput, reliability and cross-platform/cross-language support. The \uwds
API is specifically discussed hereafter, in section~\ref{api}.

\uwds is meant to broadcast complex environment representations (typically
including large geometric datasets, like meshes) in real-time. \uwds itself does
not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor
fusion, physics simulation, etc. -- are performed by the clients themselves) and
as such, the performance bottleneck is essentially the network's data
throughput.  In that regard, one of the simple yet critical optimisation
performed by \uwds is automatic caching of mesh data. Mesh data are not
transmitted when nodes are updated; only a hash value of the mesh data. The
client can then request the full data whenever they actually need it.

@@

\subsubsection{3D Visualisation and manipulation}

Interestingly, while \uwds deals with 3D geometries and scenes, it does
represent 3D entities purely as data structures: no visual representation is
involved (and as such, the \uwds server and core libraries do not depend on any
graphics library like OpenGL). However, for all practical purposes, the ability
to visualise the content of a scene is desirable. \uwds provides a standard
client, {\tt uwds view}, that performs real-time 3D rendering of worlds,
using OpenGL (Figure~\ref{fig|uwds-view}).

@@

Geometric data like meshes can be loaded into a world (or rather, into the whole
network as geometric data are treated as immutable resources shared amongst all
worlds) by simply sending an array of vertices and faces to the \uwds server.
As such, any client can load 3D data that is automatically made available to the
whole network. For example, an object recognition module recognises a new
object. It sends the mesh to the \uwds server and creates a new node which
references the mesh. At the next frame update, the 3D viewer sees a new node
referencing a yet unknown mesh. The viewer request the mesh data from the
server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
the node transformation is updated, the 3D viewer can re-render the node without
having to download the mesh from the server again.

@@

Specialised command-line tools are also available to list at run-time the worlds
and their content ({\tt uwds ls}) or to display detailed informations on a
specific node ({\tt uwds show}).

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{l2tor-photo2}
    \caption{A user manipulating objects on top of a RFID mat (providing object
    localisation) with a NAO robot, while an operator monitors
    the interaction on a tablet that display an {\tt uwds view} client.}

@@

\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
in order to conceptualise and visualise the spatial relations and visibility of
the physical objects that participants interact with. In one of the scenarios of this
project, participants are instructed by a NAO robot to manipulate a number of
Duplo-like animals on top of a printed background paper which serves as a
landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
pedagogical strategies to promote spatial relation learning by object
manipulation. Such instructions include commands like to place the elephant
\textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside}
the house.  To assess the performance, the robot and the operator need to
compute spatial relation between objects from the visual perspective of the
participant.

@@

The interaction scenario requires a number of \uwds clients. The client are
distributed across two computers (the robot and a Microsoft Windows-operated
tablet) that communicate with each other in order to synchronise the worlds
representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
the \uwds network, that consists in the following clients:

@@

Finally, we are evaluating the use of a hybrid platform that utilises both virtual and
physical objects at the same time.
